Blame and Coercion: Together Again for the Second Time

### 2019-10-15 paper.tex
normal coercion

\newcommand{\GTLC}{\texttt{GTLC+}}
\newcommand{\figref}[1]{Fig.~\ref{#1}}
\newcommand{\stxrule}[3]{#1 & ::= & #3 & \text{#2}\\}
\newcommand{\stxrulecont}[1]{& | & #1 & \\}
\newcommand{\funrule}[3]{#1 &=& #2 & #3\\}
\newcommand{\redrule}[3]{#1 & \longrightarrow_B & #2 & #3\\}
\newcommand{\redruleS}[3]{#1 & \longrightarrow_{S(C)} & #2 & #3\\}
\newcommand{\comprule}[4]{#1 & \fatsemi^\ell & #2 & = & #3 & #4 \\}
\newcommand{\plus}[0]{+}
\newcommand{\judgetype}[3]{#1 \vdash #2 : #3}
\newcommand{\lazyUD}{$\mathtt{Lazy UD}$}
\newcommand{\lazyD}{$\mathtt{Lazy D}$}
\newcommand{\eager}{$\mathtt{Eager}$}
\newcommand{\hyperCoercionI}[0]{\mathtt{id\star}}
\newcommand{\hyperCoercionC}[3]{#1 \overset{#2}{\curvearrowright} #3}
\newcommand{\sOOinspect}[3]{\mathtt{Eval} \; #1 \; #2 \; #3}
\newcommand{\sOOreturn}[2]{\mathtt{Cont} \; #2 \; #1}
\newcommand{\sOOhalt}[1]{\mathtt{Halt} \; #1}
\newcommand{\TOOdyn}[0]{\star}
\newcommand{\TOOpre}[1]{#1}
\newcommand{\POOunit}[0]{\mathtt{Unit}}
\newcommand{\POOfun}[2]{#1 \shortrightarrow #2}
\newcommand{\POOprod}[2]{#1 \times #2}
\newcommand{\POOsum}[2]{#1 \plus #2}
\newcommand{\eOOvar}[1]{#1}
\newcommand{\eOOsole}[0]{\mathtt{unit}}
\newcommand{\eOOlam}[4]{\lambda^{#1\rightarrow{}#2}#3.#4}
\newcommand{\eOOapp}[2]{#1 \; #2}
\newcommand{\eOOcons}[2]{\mathtt{cons} \; #1 \; #2}
\newcommand{\eOOcar}[1]{\mathtt{fst} \; #1}
\newcommand{\eOOcdr}[1]{\mathtt{snd} \; #1}
\newcommand{\eOOinl}[1]{\mathtt{inl} \; #1}
\newcommand{\eOOinr}[1]{\mathtt{inr} \; #1}
\newcommand{\eOOcase}[3]{\mathtt{case} \; #1 \; #2 \; #3}
\newcommand{\eOOcast}[4]{#1 \langle \cOOcast{#2}{#3}{#4} \rangle}
\newcommand{\eOOblame}[1]{\mathtt{blame} \; #1}
\newcommand{\oOOinj}{\mathtt{dyn}}
\newcommand{\oOOsole}{\mathtt{unit}}
\newcommand{\oOOfun}{\mathtt{fun}}
\newcommand{\oOOcons}{\mathtt{cons}}
\newcommand{\oOOinl}{\mathtt{inl}}
\newcommand{\oOOinr}{\mathtt{inr}}
\newcommand{\oOOblame}[1]{\mathtt{blame} \; #1}
\newcommand{\cOOcast}[3]{#1 \overset{#2}{\Rightarrow} #3}
\newcommand{\vOOcast}[2]{#1\langle#2\rangle}
\newcommand{\vOOfun}[3]{\mathtt{fun} \; #1 \; #2 \; #3}
\newcommand{\vOOtt}[0]{\mathtt{unit}}
\newcommand{\vOOcons}[2]{\mathtt{cons}\;#1\;#2}
\newcommand{\vOOinl}[1]{\mathtt{inl}\;#1}
\newcommand{\vOOinr}[1]{\mathtt{inr}\;#1}
\newcommand{\rOOsucc}[1]{\mathtt{succ}\;#1}
\newcommand{\rOOfail}[1]{\mathtt{fail}\;#1}
\newcommand{\typingHC}[3]{#1 : #2 \Longrightarrow #3}
\newcommand{\hcvOOinj}[2]{\mathtt{inj} \; #2}
\newcommand{\hcvOOfun}[5]{\mathtt{fun} \; #1 \; #2 \; #3 \; #4 \; #5}
\newcommand{\hcvOOtt}[0]{\mathtt{unit}}
\newcommand{\hcvOOcons}[4]{\mathtt{cons}\;#1\;#2\;#3\;#4}
\newcommand{\hcvOOinl}[2]{\mathtt{inl}\;#1\;#2}
\newcommand{\hcvOOinr}[2]{\mathtt{inr}\;#1\;#2}
\newcommand{\hckOOmt}[0]{\mathtt{stop}}
\newcommand{\hckOOconsI}[3]{\mathtt{cons_1}\;#1\;#2\;#3}
\newcommand{\hckOOappII}[2]{\mathtt{app_2}\;#1\;#2}
\newcommand{\sidecond}[1]{\text{if}\;#1}

@@

	Reduction \fbox{$ s \longrightarrow_{D,B} s $}
	\[
	\begin{array}{rclr}

@@

	Transitive closure of reduction \fbox{$ s \longrightarrow_{D,B}^{*} s $}\\

	Evaluation \fbox{$ e \Downarrow_{D,B} o $}
	\[
	\inference{
		\sOOinspect{e}{\emptyset}{cont(\hckOOmt)} \longrightarrow_{B}^{*} 
		\sOOhalt{o}
	}{
		e \Downarrow_{D,B} o
	}
	\]

@@

Fig.~\ref{machine-cekc} defines the dynamic semantcis of the \lazyD{} Blame 
Calculus in the style of CEK machines (\citet{felleisen1986control}). 
We choose CEK machine because to achieve space efficiency, we need to change 
continuations. Thus, abstract machines that include continuation in states is 
quite convenient for us. And CEK is one of the simplest machine of this kind.

@@

@@

\section{Background: Coercion Calculus} \label{sec:coercion-calculus}

\todo[inline]{Which version of normal coercion should this section review?}

\subsection{\lazyD{} Coercion Calculus}

\subsection{\lazyUD{} Coercion Calculus}

@@

%	Coercion typing \fbox{$ c : T \Longrightarrow T $}
%	\begin{gather*}
%	\inference{}{\typingHC{\hyperCoercionI}{\TOOdyn}{\TOOdyn}}
%	\quad
%	\inference{
%		\typingHC{h}{T_1}{P_1} &
%		\typingHC{m}{P_1}{P_2} &
%		\typingHC{t}{P_2}{T_2}
%	}{
%		\typingHC{\hyperCoercionC{h}{m}{t}}{T_1}{T_2}
%	}
%	\end{gather*}
%	
%	Head typing \fbox{$ \typingHC{h}{T}{P} $}
%	\begin{gather*}
%	\inference{}{\typingHC{\epsilon}{P}{P}}
%	\quad
%	\inference{}{\typingHC{?^l}{\TOOdyn}{P}}
%	\end{gather*}
%	
%	Middle typing \fbox{$ \typingHC{m}{P}{P} $}
%	\begin{gather*}
%	\inference{}{\typingHC{\POOunit}{\POOunit}{\POOunit}}
%	\\
%	\inference{
%		\typingHC{c_1}{T_3}{T_1} &
%		\typingHC{c_2}{T_2}{T_4}
%	}{
%		\typingHC{\POOfun{c_1}{c_2}}{\POOfun{T_1}{T_2}}{\POOfun{T_3}{T_4}}
%	}
%	\\
%	\inference{
%		\typingHC{c_1}{T_1}{T_3} &
%		\typingHC{c_2}{T_2}{T_4}
%	}{
%		\typingHC{\POOprod{c_1}{c_2}}{\POOprod{T_1}{T_2}}{\POOprod{T_3}{T_4}}
%	}
%	\\
%	\inference{
%		\typingHC{c_1}{T_1}{T_3} &
%		\typingHC{c_2}{T_2}{T_4}
%	}{
%		\typingHC{\POOsum{c_1}{c_2}}{\POOsum{T_1}{T_2}}{\POOsum{T_3}{T_4}}
%	}
%	\end{gather*}
%	
%	Tail typing \fbox{$ \typingHC{t}{P}{T} $}
%	\begin{gather*}
%	\inference{}{\typingHC{\epsilon}{P}{P}} \quad
%	\inference{}{\typingHC{!}{P}{\TOOdyn}} \quad
%	\inference{}{\typingHC{\bot^l}{P}{T}} \quad
%	\end{gather*}

>>>>

\newcommand{\GTLC}{\texttt{GTLC+}}
\newcommand{\figref}[1]{Fig.~\ref{#1}}
\newcommand{\stxrule}[3]{#1 & ::= & #3 & \text{#2}\\}
\newcommand{\stxrulecont}[1]{& | & #1 & \\}
\newcommand{\funrule}[3]{#1 &=& #2 & #3\\}
\newcommand{\redruleS}[3]{#1 & \longrightarrow_{S(C)} & #2 & #3\\}
\newcommand{\comprule}[4]{#1 & \fatsemi^\ell & #2 & = & #3 & #4 \\}
\newcommand{\plus}[0]{+}
\newcommand{\judgetype}[3]{#1 \vdash #2 : #3}
\newcommand{\lazyUD}{$\mathtt{Lazy UD}$}
\newcommand{\lazyD}{$\mathtt{Lazy D}$}
\newcommand{\eager}{$\mathtt{Eager}$}
\newcommand{\sOOinspect}[3]{\mathtt{Eval} \; #1 \; #2 \; #3}
\newcommand{\sOOreturn}[2]{\mathtt{Cont} \; #2 \; #1}
\newcommand{\sOOhalt}[1]{\mathtt{Halt} \; #1}
\newcommand{\TOOdyn}[0]{\star}
\newcommand{\TOOpre}[1]{#1}
\newcommand{\POOunit}[0]{\mathtt{Unit}}
\newcommand{\POOfun}[2]{#1 \shortrightarrow #2}
\newcommand{\POOprod}[2]{#1 \times #2}
\newcommand{\POOsum}[2]{#1 \plus #2}
\newcommand{\eOOvar}[1]{#1}
\newcommand{\eOOsole}[0]{\mathtt{unit}}
\newcommand{\eOOlam}[4]{\lambda^{#1\rightarrow{}#2}#3.#4}
\newcommand{\eOOapp}[2]{#1 \; #2}
\newcommand{\eOOcons}[2]{\mathtt{cons} \; #1 \; #2}
\newcommand{\eOOcar}[1]{\mathtt{fst} \; #1}
\newcommand{\eOOcdr}[1]{\mathtt{snd} \; #1}
\newcommand{\eOOinl}[1]{\mathtt{inl} \; #1}
\newcommand{\eOOinr}[1]{\mathtt{inr} \; #1}
\newcommand{\eOOcase}[3]{\mathtt{case} \; #1 \; #2 \; #3}
\newcommand{\eOOcast}[4]{#1 \langle \cOOcast{#2}{#3}{#4} \rangle}
\newcommand{\eOOblame}[1]{\mathtt{blame} \; #1}
\newcommand{\oOOinj}{\mathtt{dyn}}
\newcommand{\oOOsole}{\mathtt{unit}}
\newcommand{\oOOfun}{\mathtt{fun}}
\newcommand{\oOOcons}{\mathtt{cons}}
\newcommand{\oOOinl}{\mathtt{inl}}
\newcommand{\oOOinr}{\mathtt{inr}}
\newcommand{\oOOblame}[1]{\mathtt{blame} \; #1}
\newcommand{\cOOcast}[3]{#1 \overset{#2}{\Rightarrow} #3}
\newcommand{\vOOcast}[2]{#1\langle#2\rangle}
\newcommand{\vOOfun}[3]{\mathtt{fun} \; #1 \; #2 \; #3}
\newcommand{\vOOtt}[0]{\mathtt{unit}}
\newcommand{\vOOcons}[2]{\mathtt{cons}\;#1\;#2}
\newcommand{\vOOinl}[1]{\mathtt{inl}\;#1}
\newcommand{\vOOinr}[1]{\mathtt{inr}\;#1}
\newcommand{\rOOsucc}[1]{\mathtt{succ}\;#1}
\newcommand{\rOOfail}[1]{\mathtt{fail}\;#1}
\newcommand{\typingHC}[3]{#1 : #2 \Longrightarrow #3}
\newcommand{\hcvOOinj}[2]{\mathtt{inj} \; #2}
\newcommand{\hcvOOfun}[5]{\mathtt{fun} \; #1 \; #2 \; #3 \; #4 \; #5}
\newcommand{\hcvOOtt}[0]{\mathtt{unit}}
\newcommand{\hcvOOcons}[4]{\mathtt{cons}\;#1\;#2\;#3\;#4}
\newcommand{\hcvOOinl}[2]{\mathtt{inl}\;#1\;#2}
\newcommand{\hcvOOinr}[2]{\mathtt{inr}\;#1\;#2}
\newcommand{\hckOOmt}[0]{\mathtt{stop}}
\newcommand{\hckOOconsI}[3]{\mathtt{cons_1}\;#1\;#2\;#3}
\newcommand{\hckOOappII}[2]{\mathtt{app_2}\;#1\;#2}
\newcommand{\sidecond}[1]{\text{if}\;#1}
% Lazy D blame calculus
\newcommand{\judgeDBreduce}[2]{#1 \longrightarrow_{\mathcal{D}} #2}
\newcommand{\judgeDBreduceTrans}[2]{#1 \longrightarrow_{\mathcal{D}}^{*} #2}
\newcommand{\judgeDBeval}[2]{#1 \Downarrow_{\mathcal{D}} #2}
\newcommand{\redrule}[3]{#1 & \longrightarrow_\mathcal{D} & #2 & #3\\}
% Normal Coercion
\newcommand{\ncProj}[2]{#1?^{#2}}
\newcommand{\ncInj}[1]{#1!}
\newcommand{\ncId}[0]{\iota}
\newcommand{\ncSeq}[2]{#1;#2}
\newcommand{\ncFail}[1]{\bot^{#1}}
\newcommand{\ncFun}[2]{\POOfun{#1}{#2}}
\newcommand{\ncProd}[2]{\POOprod{#1}{#2}}
\newcommand{\ncSum}[2]{\POOsum{#1}{#2}}
% Hyper-coercion
\newcommand{\hyperCoercionI}[0]{\mathtt{id\star}}
\newcommand{\hyperCoercionC}[3]{#1 \overset{#2}{\curvearrowright} #3}

@@

	Reduction \fbox{$ \judgeDBreduce{s}{s} $}
	\[
	\begin{array}{rclr}

@@

%	Transitive closure of reduction \fbox{$ s \longrightarrow_{D,B}^{*} s $}\\

	Evaluation \fbox{$\judgeDBeval{e}{o}$}
	\[
	\inference{
		\sOOinspect{e}{\emptyset}{cont(\hckOOmt)} \longrightarrow_{B}^{*} 
		\sOOhalt{o}
	}{
		\judgeDBeval{e}{o}
	}
	\]

@@

Fig.~\ref{machine-cekc} defines the dynamic semantics of the \lazyD{} Blame 
Calculus in the style of CEK machines (\citet{felleisen1986control}). 
We conjecture that this CEK machines agree with the \lazyD{} blame calculus in 
\citet{siek2012interpretations}.
We will show in \hyperref{sec:hyper-coercion} that this abstract machine is 
equivalent to an abstract machine that represents casts with hyper-coercions.

@@

Transitive closure of reduction ($ \judgeDBreduceTrans{s}{s} $) and evaluation 
are 
standard.

@@

\section{Background: Normal Coercion} \label{sec:coercion-calculus}

In this section we review the normal coercion (NC) in 
\citet{siek2012interpretations}. We ignore sum types and product types in this 
section, because they were not presented in the paper. We assume a basic 
familiarity with coercion, and direct readers who do not know it to 
\citet{siek2012interpretations}.

As the name suggests, 
hyper-coercion is largely inspired by coercion, especially its subst NC. We 
will highlight the similarity of normal coercion and hyper-coercion, and hope 
this would convince you that hyper-coercion should be as difficult to implement 
as NC.

\begin{figure}
	\[
	\begin{array}{rclr}
	\stxrule{I}{injectable types (\lazyD)}{
		\POOfun{T}{T}}
	\stxrule{I}{injectable types (\lazyUD)}{
		\POOfun{\TOOdyn}{\TOOdyn}
	}
	\stxrule{c}{coercions}{
		\ncInj{I} \mid
		\ncProj{I}{l} \mid
		\ncId \mid
		\ncFail{l} \mid
		\ncSeq{c}{c} \mid
		\ncFun{c}{c}
	}
	\stxrule{\bar{c}}{wrapper coercions}{	
		\ncInj{I} \mid
		\ncFun{\hat{c}}{\hat{c}} \mid
		\ncSeq{\ncFun{\hat{c}}{\hat{c}}}{\ncInj{I}}
	}
%	\stxrulecont{
%		\ncProd{\hat{c}}{\hat{c}} \mid
%		\ncSeq{\ncProd{\hat{c}}{\hat{c}}}{\ncInj{I}} \mid
%		\ncSum{\hat{c}}{\hat{c}} \mid
%		\ncSeq{\ncSum{\hat{c}}{\hat{c}}}{\ncInj{I}} \mid	
%	}
	\stxrule{\hat{c}}{normal coercions}{
		\bar{c} \mid
		\ncId \mid
		\ncFail{l} \mid
		\ncProj{I}{l} \mid
		\ncSeq{\ncProj{I}{l}}{\ncFail{l}} \mid
		\ncSeq{\ncProj{I}{l}}{\ncInj{I}}
	}
	\stxrulecont{
		\ncSeq{\ncProj{I}{l}}{\ncFun{\hat{c}}{\hat{c}}} \mid
		\ncSeq{\ncProj{I}{l}}{\ncSeq{\ncFun{\hat{c}}{\hat{c}}}{\ncInj{I}}}
	}
	\end{array}
	\]
	\caption{Syntax of Normal Coercion (NC)}
	\label{fig:normal-coercion}
\end{figure}

\figref{fig:normal-coercion} defines the normal coercion. 
Let $ I $ ranges over injectable types. An injectable type is a type that can 
be cast directly to (injection) and from (projection) $ \TOOdyn $. The 
definition of injectable type depends on blame strategy. With \lazyD, all 
pre-types are injectable. With \lazyUD, only $ \POOfun{\TOOdyn}{\TOOdyn} $ 
is injectable. 
Let $ c $ ranges over coercions, a coercion is either injection, projection, 
identity, failure, sequencing, or function coercion.
Let $ \bar{c} $ ranges over wrapper coercions, which represent casts in $ 
\vOOcast{v}{c} $.
Let $ \hat{c} $ ranges over normal coercions. If we inline $ \bar{c} $ and 
re-arrange cases a little bit, the definition becomes:
\[
\begin{array}{rclr}
\stxrule{\hat{c}}{normal coercions}{
	\ncFun{\hat{c}}{\hat{c}} \mid
	\ncSeq{\ncProj{I}{l}}{\ncFun{\hat{c}}{\hat{c}}} \mid
	\ncSeq{\ncFun{\hat{c}}{\hat{c}}}{\ncInj{I}} \mid
	\ncSeq{\ncProj{I}{l}}{\ncSeq{\ncFun{\hat{c}}{\hat{c}}}{\ncInj{I}}} \mid
	\ncSeq{\ncProj{I}{l}}{\ncFail{l}}
}
\stxrulecont{
	\ncId \mid
	\ncProj{I}{l} \mid
	\ncInj{I} \mid
	\ncSeq{\ncProj{I}{l}}{\ncInj{I}} \mid
	\ncFail{l} \mid
}
\end{array}
\]

Three observations on the definition leads to hyper-coercion: 
\begin{enumerate}
	\item The length of normal coercion is at most three.
	\item Projections are always at the beginning when presented.
	\item Injections and failures are always at the end when presented.
\end{enumerate}

@@

	%	Coercion typing \fbox{$ c : T \Longrightarrow T $}
	%	\begin{gather*}
	%	\inference{}{\typingHC{\hyperCoercionI}{\TOOdyn}{\TOOdyn}}
	%	\quad
	%	\inference{
	%		\typingHC{h}{T_1}{P_1} &
	%		\typingHC{m}{P_1}{P_2} &
	%		\typingHC{t}{P_2}{T_2}
	%	}{
	%		\typingHC{\hyperCoercionC{h}{m}{t}}{T_1}{T_2}
	%	}
	%	\end{gather*}
	%	
	%	Head typing \fbox{$ \typingHC{h}{T}{P} $}
	%	\begin{gather*}
	%	\inference{}{\typingHC{\epsilon}{P}{P}}
	%	\quad
	%	\inference{}{\typingHC{?^l}{\TOOdyn}{P}}
	%	\end{gather*}
	%	
	%	Middle typing \fbox{$ \typingHC{m}{P}{P} $}
	%	\begin{gather*}
	%	\inference{}{\typingHC{\POOunit}{\POOunit}{\POOunit}}
	%	\\
	%	\inference{
	%		\typingHC{c_1}{T_3}{T_1} &
	%		\typingHC{c_2}{T_2}{T_4}
	%	}{
	%		\typingHC{\POOfun{c_1}{c_2}}{\POOfun{T_1}{T_2}}{\POOfun{T_3}{T_4}}
	%	}
	%	\\
	%	\inference{
	%		\typingHC{c_1}{T_1}{T_3} &
	%		\typingHC{c_2}{T_2}{T_4}
	%	}{
	%		
	%\typingHC{\POOprod{c_1}{c_2}}{\POOprod{T_1}{T_2}}{\POOprod{T_3}{T_4}}
	%	}
	%	\\
	%	\inference{
	%		\typingHC{c_1}{T_1}{T_3} &
	%		\typingHC{c_2}{T_2}{T_4}
	%	}{
	%		\typingHC{\POOsum{c_1}{c_2}}{\POOsum{T_1}{T_2}}{\POOsum{T_3}{T_4}}
	%	}
	%	\end{gather*}
	%	
	%	Tail typing \fbox{$ \typingHC{t}{P}{T} $}
	%	\begin{gather*}
	%	\inference{}{\typingHC{\epsilon}{P}{P}} \quad
	%	\inference{}{\typingHC{!}{P}{\TOOdyn}} \quad
	%	\inference{}{\typingHC{\bot^l}{P}{T}} \quad
	%	\end{gather*}
